---
title: "Assignment3_IT_Helpdesk_Modeling"
author: "Minghao Du"
output:
  word_document: default
  html_notebook: default
---

# Set-up
Load packages. 
```{r}
rm(list = ls())
library(readxl)
library(tidyverse)
library(ggplot2)
```

For this and the next assignments (Assignment 3 and 4), you will use the dataset from https://www.ibm.com/communities/analytics/watson-analytics-blog/it-help-desk/. (the link is now invalid.) 
This data is from an Information Technology (IT) department interested in improving the satisfaction of customers.  
To start their analysis, they constructed this data set of 100,000 closed tickets that were filed at their help desk. 

Load the dataset and save it as ithelp. 
```{r}
ithelp <- read_excel("WA_Fn-UseC_-IT-Help-Desk.xlsx")
```

```{r}
glimpse(ithelp)
```

A total of 100,000 rows. Each row represent an individual request. The data includes:

* Requestor: employee who submitted the ticket
* RequestorSeniority: employee’s seniority within the company
* ITOwner: IT employee who serviced the ticket
* FileAgainst: functional area against which the ticket was filed (systems, software, hardware, access)
* TicketType: whether the ticket was a request for new services or an issue with existing services
* Severity: submitter-assigned severity of the ticket
* Priority: IT-assigned priority of the ticket
* daysOpen: number of days the ticket was open
* Satisfaction: satisfaction with the resolution of the ticket (reported by the submitter)

```{r}
summary(ithelp)
```

Some 'character' variables behave unexpectedly. Let's convert all character variables into factors, which indicate categorical variables in R. 

```{r}
ithelp<-ithelp%>%
  mutate(ticket=as.factor(ticket),
         Requestor=as.factor(Requestor),
         RequestorSeniority=as.factor(RequestorSeniority),
         ITOwner=as.factor(ITOwner),
         FiledAgainst=as.factor(FiledAgainst),
         TicketType=as.factor(TicketType),
         Severity=as.factor(Severity),
         Priority=as.factor(Priority),
         Satisfaction=as.factor(Satisfaction))
```

Let's check the summary again. 
```{r}
summary(ithelp)
```

It turns out that there are many cases with "Unknown" Satisfaction. Let's exclude these cases from the analysis (Step 1: filter). Next, we may build a multi-class classification model (Unsatisfied, Satisfied, Highly satisfied), but let's simplify it to a binary classifier and identify "Unsatisfied" cases, which are problematic. Create a new variable, "negative", which indicates if a user's feedback is negative (Step 2: mutate). The following code will do these jobs for you. 

```{r}
ithelp<-ithelp%>%
  filter(Satisfaction!="0 - Unknown")%>%
  mutate(negative=as.factor(ifelse(Satisfaction =="1 - Unsatisfied","Yes","No")))

summary(ithelp)
```

# Q1. Classification tree model
```{r}
library(rpart)
library(rpart.plot)
```

Build a model to predict "negative" with the following variables. Fill in the blank to complete the code below. 
* RequestorSeniority
* FiledAgainst
* TicketType
* Severity
* Priority
* daysOpen

Set the control as `control=rpart.control(cp=0.001)`. 
```{r}
ct_model<-rpart(negative ~ RequestorSeniority+FiledAgainst+TicketType+Severity+Priority+daysOpen,    # model formula
                data = ithelp,             # dataset
                method="class",            # "class" indicates a classification tree model 
                control=rpart.control(cp=0.001))   # tree control parameters. 
```

Display the tree model with `rpart.plot` and answer the following questions based on this model result.
```{r}
rpart.plot(ct_model)

```

Question 1: Which one of the following characterizes the case that receive negative feedback (i.e. the probability of negative feedback greater than 50%)? 

(1) Not closed (i.e. resolved) within 12 days (daysOpen=>13 days) 
(2) Priority is High
(3) Not filed against Hardware (i.e. filed against Access/Login, Software, or Systems)
(4) Severity is either minor or normal
(5) All conditions above should be met. 


Question 2: What is the probability that a case receives negative feedback if the case is resolved within 12 days? 
(1) 27%
(2) 41%
(3) 80%
(4) Cannot know.

# Logit Regression Model
Build a logit regression model to predict "negative" with the following variables. Fill in the blank to complete the code below. 
* RequestorSeniority
* FiledAgainst
* TicketType
* Severity
* Priority
* daysOpen

```{r}
logit_model<-glm(negative ~ RequestorSeniority+FiledAgainst+TicketType+Severity+Priority+daysOpen,             
                 family="binomial",
                 data = ithelp)
summary(logit_model)
```
Question 3: Which of the following is NOT an implication that you can draw from this logit regression result? 

(1) An increase in the daysOpen is associated with a higher chance of negative feedback.
(2) A request with high priority is more likely to receive negative feedback.
(3) Compared to Junior, a request from Senior is more likely to receive negative feedback.
(4) All of the above


Question 4. Save the predicted probability from a logit model by creating a new variable *logit_pred* in the dataset. What is the probability that the ticket number 1404 receives negative feedback? 
```{r}
ithelp$logit_pred<-predict(logit_model, type = "response")
ithelp%>%
  filter(ticket == 1404)
```

(1) 18%
(2) 23%
(3) 53%
(4) 118%

Question 5. The same as the previous assignments, your last task is creating a report. Change the author name on the top of this R markdown file to yours. Compile this R markdown file into a Word or html document and submit it through the course Blackboard. Your report should contain all codes and results of in-class exercises as well as the assignment questions and answers. 




